The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-11-26 conf_article/neon.tex
adopted pdf from remote origin
Minor fixes

Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible. On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance. On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279. Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon. These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above. Moreover, the tools from ~\ref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that does not have to be modifications of Muon.

@@

We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}. First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$. We record the accuracy after 8 epochs of training (\cref{fig:muon_alphas}).

@@

We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon on CIFAR-airbench. To preserve the algorithms, we do not normalize the weights of the network each iteration. The parameters are in the table.

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|c|c|}
	  \hline
	  Method & lr  & momentum   & val\_accuracy, \% \\
	  \hline
	  NSGD & ? & ? & ? \\
	  \hdashline
	  SignSGD & ? $\cdot$ ? & ? & ? \\
	  \hdashline
	  Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
	  \hdashline
	  F-Muon & 0.40 & 0.6 & $94.02 \pm 0.13$\\
	  \hdashline
	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
	  \hline
	\end{tabular}
  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
  \label{tbl:cifar_lrs}
  \end{table}

>>>>

Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible. On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.01 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance. On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279. Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon. These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above. Moreover, the tools from ~\ref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that does not have to be modifications of Muon.

@@

We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}. First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$. We record the accuracy after 8 epochs of training (\cref{fig:muon_alphas}).

@@

We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon on CIFAR-airbench. NSGD and SignSGD are not heavily tuned. To preserve the algorithms, we do not normalize the weights of the network each iteration. The parameters are in the table.

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|c|c|}
	  \hline
	  Method & \texttt{lr}  & \texttt{momentum}  & val\_accuracy, \% \\
	  \hline
	  NSGD & 0.5 & 0.95 & $91.6 \pm 0.52$ \\
	  \hdashline
	  SignSGD & 0.003 & 0.95 & $91.54 \pm 0.26$ \\
	  \hdashline
	  Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
	  \hdashline
	  F-Muon & 0.40 & 0.6 & $94.01 \pm 0.13$\\
	  \hdashline
	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
	  \hline
	\end{tabular}
  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
  \label{tbl:cifar_lrs}
  \end{table}
